---
title: "CHAVA_DGE_Analysis"
author: "Carolina"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, tidy.opts=list(width.cutoff=60), tidy=TRUE) 
```
#Set Up

load required packages
```{r, message=FALSE, results='hide'}
library(DESeq2)
library(readxl)
library(dplyr)
library(ggplot2)
library(ggfortify)
library(plotly)
library(BiocManager)
library(EnhancedVolcano)
library("AnnotationDbi")
library("org.Hs.eg.db")
library(EnhancedVolcano)
library(ggvenn)
library(pheatmap)
library(clusterProfiler)
library(enrichplot)
library("biomaRt")
library("tibble")
library(PoiClaClu)
library(pheatmap)
library(RColorBrewer)

library(GGally)
library(PCAtools)
library("enrichR")
library(ggpubr)
```


set up count and condition files

```{r, echo=FALSE}
CHAVAcondition=read_excel("CHAVAcondition.xlsx")
CHAVAcounts=read.delim("CHAVAcounts.txt",row.names = 1)
CHAVAcountsonly=CHAVAcounts[-c(1:5)]

CHAVAinfo=CHAVAcondition
CHAVAinfo$Infection=as.factor(ifelse(CHAVAinfo$Infection==1,"Pos","Neg"))
CHAVAinfo$Sex=as.factor(ifelse(CHAVAinfo$Sex==1,"F","M"))
CHAVAinfo$ECG=as.factor(CHAVAinfo$ECG)
CHAVAinfo$Batch=as.factor(CHAVAinfo$Batch)
CHAVAinfo$OldECG=as.factor(CHAVAinfo$OldECG)
CHAVAinfo$Cat=as.factor(CHAVAinfo$Cat)
CHAVAinfo$Origin=as.factor(CHAVAinfo$Origin)
CHAVAinfo$Batch=as.factor(CHAVAinfo$Batch)
CHAVAinfo$RIN=as.factor(round(CHAVAinfo$RIN))
CHAVAinfo$AgeRange=cut(CHAVAinfo$Age,breaks=3)
CHAVAinfo$AgeCategory=CHAVAinfo$AgeRange
levels(CHAVAinfo$AgeCategory)=c("Young", "Mid","Old","Oldest")

#here i subset for patients that have values for the ECG, and I remove B140 since incorrect sex and idk what else might be wrong
ECGs=subset(CHAVAinfo,ECG!="NA")$NovogeneCode
CHAVA_ECG=CHAVAinfo[CHAVAinfo$NovogeneCode %in% c(ECGs),]
CHAVA_ECG=droplevels(CHAVA_ECG)
Counts_ECG=subset(CHAVAcountsonly, select=ECGs)
```

Make demographics table
```{r}
CHAVA_table=CreateTableOne(data=subset(CHAVA_ECG), strata="Infection", vars=c("Age","Sex","Origin","Cat"))

print(CHAVA_table, exact = c("Origin","Sex","Cat"))
Neg_CHAVA_table=CreateTableOne(data=subset(CHAVA_ECG,Infection=="Neg" & Cat != "A2"), strata="Cat", vars=c("Age","Sex","Origin"))
print(Neg_CHAVA_table, exact = c("Origin","Sex"))


Pos_CHAVA_table=CreateTableOne(data=subset(CHAVA_ECG,Infection=="Pos" & Cat != "A2"), strata="Cat", vars=c("Age","Sex","Origin"))
print(Pos_CHAVA_table, exact = c("Origin","Sex"))
```

remove rRNA counts
```{r}
ensembl=useMart("ENSEMBL_MART_ENSEMBL", host = "https://www.ensembl.org")
ensembl = useDataset("hsapiens_gene_ensembl", mart=ensembl)
rRNAs=biomaRt::getBM(values="rRNA", 
               filters="biotype", 
               attributes=c("ensembl_gene_id", "external_gene_name", "gene_biotype"), 
               mart = ensembl)
Counts_ECG=Counts_ECG[!(row.names(Counts_ECG) %in% rRNAs$ensembl_gene_id),]
```

## PCA

set up a dds object with all possible variable to explore and filter out low-count genes
```{r}
dds_Var=DESeqDataSetFromMatrix(countData = Counts_ECG, colData =CHAVA_ECG, design= ~Sex + Batch + Origin + Cat+ AgeCategory +  Infection)

keepddsVar= rowSums(counts(dds_Var)>= 10) >=3 #keeps genes with at least 3 samples with 10 counts
dds_Var <- dds_Var[keepddsVar,]
```

vst transform to meet assumptions for PCA analysis. 
```{r}
vst_Var <- vst(dds_Var, blind=FALSE)
```
```{r, message = FALSE}
rv <- rowVars(assay(vst_Var))
pcagenes <-  order(rv, decreasing=TRUE)[seq_len(min(1000, length(rv)))]
pca <- prcomp(t(assay(vst_Var)[pcagenes,]))
percentVar <- pca$sdev^2 / sum( pca$sdev^2 )

autoplot(pca, data=CHAVA_ECG, colour="Sex", size=4) + theme_classic()
ggsave("SexPCA.pdf")


autoplot(pca, data=CHAVA_ECG, colour="Batch", size=4) + theme_classic() + scale_color_manual(values=c("purple","darkgrey"))
ggsave("BatchEffect.pdf")


ggpairs(pca$x, columns = 1:5, aes(colour=CHAVA_ECG$Infection), upper =list(continuous= "blank"),diag = list(continuous = wrap("densityDiag",alpha=0.5)),legend=c(2,1),lower = list(continuous = wrap("points", alpha = 0.9, size=0.7))) + 
  scale_color_manual(values=c("#4349e8","#f5474a"))+ 
  scale_fill_manual(values=c("#4349e8","#f5474a"))+ theme_bw() + labs(colour="Infection Status")
ggsave("Infection.pdf")

ggpairs(pca$x, columns = 1:5, aes(colour=CHAVA_ECG$Origin), upper =list(continuous= "blank"),diag = list(continuous = wrap("densityDiag",alpha=0.5)),legend=c(2,1),lower = list(continuous = wrap("points", alpha = 0.9, size=0.7))) + 
  scale_color_manual(values=c("#DE8333","#1c9470")) +
  scale_fill_manual(values=c("#DE8333","#1c9470")) +theme_bw() + labs(colour="Origin")
ggsave("Origin.pdf")

ggpairs(pca$x, columns = 1:5, aes(colour=CHAVA_ECG$Cat), upper =list(continuous= "blank"),diag = list(continuous = wrap("densityDiag",alpha=0.5)),legend=c(2,1),lower = list(continuous = wrap("points", alpha = 0.9, size=0.7))) + 
  scale_color_manual(values=c("#0fbab7","#1b804f","#6e016b")) +
  scale_fill_manual(values=c("#0fbab7","#1b804f","#6e016b")) +theme_bw() +  labs(colour="Heart Failure")
ggsave("AHA_Category.pdf")

ggpairs(pca$x, columns = 1:5, aes(colour=CHAVA_ECG$AgeCategory), upper =list(continuous = "blank"),diag = list(continuous = wrap("densityDiag",alpha=0.5)),legend=c(2,1),lower = list(continuous = wrap("points", alpha = 0.9, size=0.7))) + theme_bw() + labs(colour="AgeCategory")
ggsave("AgeCategory.pdf")


phenoCols=list(Infection=c(Neg="#4349e8",Pos="#f5474a"),Origin=c(CentralAm="#DE8333",Bolivia="#1c9470"),Cat=c(A="#0fbab7",B="#6e016b"))

```

# A vs B by infection status (excluding A2)

## Postive
```{r}
CHAVA_ECG_AB_Pos=subset(CHAVA_ECG, Cat!="A2"&Infection=="Pos")
CHAVA_ECG_AB_Pos=droplevels(CHAVA_ECG_AB_Pos)
Counts_ECG_AB_Pos=subset(Counts_ECG, select=CHAVA_ECG_AB_Pos$NovogeneCode)
```

### DGE
```{r}

ddsAB_Pos=DESeqDataSetFromMatrix(countData = Counts_ECG_AB_Pos, colData =CHAVA_ECG_AB_Pos, design= ~ Batch + Sex +  Cat)
keepddsAB_Pos= rowSums(counts(ddsAB_Pos)>= 10) >=3 #keeps genes with at least 3 samples with 10 counts
ddsAB_Pos <- ddsAB_Pos[keepddsAB_Pos,]
ddsAB_Pos=DESeq(ddsAB_Pos) #normalizes, estimates dispersion and fits glm

result_AB_Pos=results(ddsAB_Pos,alpha = 0.1,contrast=c("Cat", "B", "A")) 
summary(result_AB_Pos)
Sig_AB_Pos=subset(result_AB_Pos, padj < 0.1)
```

annotate results
```{r}

annot_AB_Pos <- getBM(attributes=c('ensembl_gene_id', 'external_gene_name'), 
               filters = "ensembl_gene_id", 
               values = rownames(result_AB_Pos), 
               mart = ensembl)

annot_AB_Pos<- as.data.frame(result_AB_Pos) %>% 
  rownames_to_column("ensembl_gene_id") %>% 
  left_join(annot_AB_Pos, "ensembl_gene_id") %>% 
  rename(log2FC=log2FoldChange, FDR=padj)


SigAnnot_AB_Pos = subset(annot_AB_Pos, FDR < 0.1)
SigAnnot_AB_Pos[order(SigAnnot_AB_Pos$log2FC), ]

#write.csv(SigAnnot_AB_Pos,"All_AB_Posv.csv")
```
### Histogram
```{r}
vst_AB_Pos=vst(ddsAB_Pos, blind=FALSE)

topgenes_AB_Pos=arrange(annot_AB_Pos, FDR)
topgenes_AB_Pos=na.omit(topgenes_AB_Pos)
topgenes_AB_Pos[topgenes_AB_Pos==""]="No GeneSymb"
sigcount=sum(topgenes_AB_Pos$FDR < 0.1)
if (sigcount <=2){sigcount=2}
topgenesnames_AB_Pos <- topgenes_AB_Pos$external_gene_name[1:sigcount]
topgenes_AB_Pos <- topgenes_AB_Pos$ensembl_gene_id[1:sigcount]

sampleinfo_AB_Pos= arrange(data.frame(colData(vst_AB_Pos)[,c("Sex","Cat")]), Cat)
sampleinfo_AB_Pos=subset(sampleinfo_AB_Pos, select= -Sex)
mat_AB_Pos= assay(vst_AB_Pos)[topgenes_AB_Pos,] - rowMeans(assay(vst_AB_Pos)[topgenes_AB_Pos,])

mat_AB_Pos = mat_AB_Pos[, rownames(sampleinfo_AB_Pos)]

pdf(file="Pos_AvB_heatmpa_0.1.pdf")
pheatmap(mat_AB_Pos, 
         annotation_col = sampleinfo_AB_Pos, 
         show_colnames = FALSE, 
         labels_row = topgenesnames_AB_Pos, 
         annotation_colors= phenoCols,
         cluster_cols=FALSE,
         color = colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100))
dev.off()
```


### volcano plot
```{r}
y_AB_Pos=-log10(min(SigAnnot_AB_Pos$FDR))+1
xmin_AB_Pos=min(SigAnnot_AB_Pos$log2FC)-1
xmax_AB_Pos=max(SigAnnot_AB_Pos$log2FC)+1

keyvals <- ifelse(
    annot_AB_Pos$log2FC < -0.58 & annot_AB_Pos$FDR<0.1, 'royalblue',
      ifelse(annot_AB_Pos$log2FC > 0.58 & annot_AB_Pos$FDR<0.1, 'red',
            ifelse((annot_AB_Pos$log2FC< -0.58 | annot_AB_Pos$log2FC> 0.58 )& annot_AB_Pos$FDR>0.1,'black',
                   ifelse(annot_AB_Pos$log2FC > -0.58 & annot_AB_Pos$log2FC < 0.58 & annot_AB_Pos$FDR<0.1,'black',
        'grey'))))

 
 
keyvals[is.na(keyvals)] <- 'grey'
names(keyvals)[keyvals == 'red'] <- 'upregulated'
names(keyvals)[keyvals == 'grey'] <- 'nonsignificant'
names(keyvals)[keyvals == 'black'] <- 'mid'
names(keyvals)[keyvals == 'royalblue'] <- 'downregulated'

EnhancedVolcano(annot_AB_Pos,
                lab = annot_AB_Pos$external_gene_name,
                x = 'log2FC',
                y = 'FDR',
                colCustom = keyvals,
                ylab = bquote(~-Log[10] ~ "FDR"),
                ylim = c(0, y_AB_Pos),
                xlim = c(xmin_AB_Pos,xmax_AB_Pos),
                title = 'B pos vs A Pos',
                subtitle = NULL,
                legendPosition = "top",
                boxedLabels = TRUE,
                drawConnectors = TRUE,
                legendLabSize = 10,
                legendIconSize = 3,
                axisLabSize = 10,
                titleLabSize = 15,
                shadeAlpha = 0.5,
                colAlpha = 0.7,
                pCutoff = 0.1,
                FCcutoff = 0.58,
                pointSize = 1.5,
                labSize = 5.0)

ggsave("Pos_AvB_Volcano.pdf", width=8)
```


## Negative
```{r}
CHAVA_ECG_AB_Neg=subset(CHAVA_ECG, Cat!="A2"&Infection=="Neg")
CHAVA_ECG_AB_Neg=droplevels(CHAVA_ECG_AB_Neg)
Counts_ECG_AB_Neg=subset(Counts_ECG, select=CHAVA_ECG_AB_Neg$NovogeneCode)
```

### DEG
```{r}
ddsAB_Neg=DESeqDataSetFromMatrix(countData = Counts_ECG_AB_Neg, colData =CHAVA_ECG_AB_Neg, design= ~Batch +Sex  + Cat)
keepddsAB_Neg= rowSums(counts(ddsAB_Neg)>= 10) >=3 #keeps genes with at least 3 samples with 10 counts
ddsAB_Neg <- ddsAB_Neg[keepddsAB_Neg,]
ddsAB_Neg=DESeq(ddsAB_Neg) #normalizes, estimates dispersion and fits glm

result_AB_Neg=results(ddsAB_Neg,alpha = 0.1,contrast=c("Cat", "B", "A")) 
summary(result_AB_Neg)
Sig_AB_Neg=subset(result_AB_Neg, padj < 0.1)

```
annotate results
```{r}

annot_AB_Neg <- getBM(attributes=c('ensembl_gene_id', 'external_gene_name'), 
               filters = "ensembl_gene_id", 
               values = rownames(result_AB_Neg), 
               mart = ensembl)

annot_AB_Neg<- as.data.frame(result_AB_Neg) %>% 
  rownames_to_column("ensembl_gene_id") %>% 
  left_join(annot_AB_Neg, "ensembl_gene_id") %>% 
  rename(log2FC=log2FoldChange, FDR=padj)

SigAnnot_AB_Neg = subset(annot_AB_Neg, FDR < 0.1)

write.csv(SigAnnot_AB_Neg,"All_AB_Neg.csv")
```
### heatmap
```{r}
vst_AB_Neg=vst(ddsAB_Neg, blind=FALSE)

topgenes_AB_Neg=arrange(annot_AB_Neg, FDR)
topgenes_AB_Neg=na.omit(topgenes_AB_Neg)
topgenes_AB_Neg[topgenes_AB_Neg==""]="No GeneSymb"
sigcount=sum(topgenes_AB_Neg$FDR < 0.1)
if (sigcount <=2){sigcount=2}
topgenesnames_AB_Neg <- topgenes_AB_Neg$external_gene_name[1:sigcount]
topgenes_AB_Neg <- topgenes_AB_Neg$ensembl_gene_id[1:sigcount]

sampleinfo_AB_Neg= arrange(data.frame(colData(vst_AB_Neg)[,c("Sex","Cat")]), Cat)
sampleinfo_AB_Neg=subset(sampleinfo_AB_Neg, select= -Sex)
mat_AB_Neg= assay(vst_AB_Neg)[topgenes_AB_Neg,] - rowMeans(assay(vst_AB_Neg)[topgenes_AB_Neg,])

mat_AB_Neg = mat_AB_Neg[, rownames(sampleinfo_AB_Neg)]

pdf(file="Neg_BvA_heatmap_0.1_ordered.pdf")
pheatmap(mat_AB_Neg, 
         annotation_col = sampleinfo_AB_Neg, 
         show_colnames = FALSE, 
         labels_row = topgenesnames_AB_Neg, 
         annotation_colors = phenoCols, 
         cluster_cols=FALSE,
         fontsize = 8,
         color = colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100))
dev.off()
```


### volcano plot
```{r}
y_AB_Neg=-log10(min(SigAnnot_AB_Neg$FDR))+1
xmin_AB_Neg=min(SigAnnot_AB_Neg$log2FC)-1
xmax_AB_Neg=max(SigAnnot_AB_Neg$log2FC)+1

keyvalsNeg <- ifelse(
    annot_AB_Neg$log2FC < -0.58 & annot_AB_Neg$FDR<0.1, 'royalblue',
      ifelse(annot_AB_Neg$log2FC > 0.58 & annot_AB_Neg$FDR<0.1, 'red',
            ifelse((annot_AB_Neg$log2FC< -0.58 | annot_AB_Neg$log2FC> 0.58 )& annot_AB_Neg$FDR>0.1,'black',
                   ifelse(annot_AB_Neg$log2FC > -0.58 & annot_AB_Neg$log2FC < 0.58 & annot_AB_Neg$FDR<0.1,'black',
        'grey'))))

 
 
keyvalsNeg[is.na(keyvalsNeg)] <- 'grey'
names(keyvalsNeg)[keyvalsNeg == 'red'] <- 'upregulated'
names(keyvalsNeg)[keyvalsNeg == 'grey'] <- 'nonsignificant'
names(keyvalsNeg)[keyvalsNeg == 'black'] <- 'mid'
names(keyvalsNeg)[keyvalsNeg == 'royalblue'] <- 'downregulated'

EnhancedVolcano(annot_AB_Neg,
                lab = annot_AB_Neg$external_gene_name,
                colCustom = keyvalsNeg,
                drawConnectors = TRUE,
                boxedLabels = TRUE,
                x = 'log2FC',
                y = 'FDR',
                ylab = bquote(~-Log[10] ~ "FDR"),
                ylim = c(0, y_AB_Neg),
                xlim = c(xmin_AB_Neg,xmax_AB_Neg),
                title = 'B Neg vs A Neg',
                subtitle = NULL,
                legendPosition = "top",
                legendLabSize = 10,
                legendIconSize = 3,
                axisLabSize = 10,
                titleLabSize = 15,
                shadeAlpha = 0.5,
                colAlpha = 0.7,
                pCutoff = 0.1,
                FCcutoff = 0.58,
                pointSize = 1.5,
                labSize = 5.0)
ggsave("Neg_BvA_volcano.pdf",width=7, height=6)
```

## DEGs in positive but not negative

```{r}
All_Genes_AB_PosNeg=list(
  `Pos_AvsB`=subset(SigAnnot_AB_Pos,log2FC < -0.58|log2FC > 0.58)$ensembl_gene_id,
  `Neg_AvsB`=subset(SigAnnot_AB_Neg,log2FC < -0.58|log2FC > 0.58)$ensembl_gene_id)

ggvenn(
  All_Genes_AB_PosNeg,
  fill_color = c("#f5474a", "#4349e8"),
  stroke_size = 0.5, set_name_size = 4)

ggsave("venn_AB_negpos.pdf")

```

FC of positive DEGs against the FC of those genes in the neg group
```{r}
FC_Table_AB= data.frame(
  `ensembl_gene_id`=subset(SigAnnot_AB_Pos,log2FC < -0.58|log2FC > 0.58)$ensembl_gene_id,
  `external_gene_name`=subset(SigAnnot_AB_Pos,log2FC < -0.58|log2FC > 0.58)$external_gene_name,
  `PosLog2FC`=subset(SigAnnot_AB_Pos,log2FC < -0.58|log2FC > 0.58)$log2FC)

test <- annot_AB_Neg %>%
   dplyr::select(ensembl_gene_id,log2FC)

FC_Table_AB <- left_join(FC_Table_AB, test, by = "ensembl_gene_id") %>% 
   dplyr::select(ensembl_gene_id,external_gene_name,PosLog2FC,log2FC) %>%
   rename(NegLog2FC=log2FC)


ggplot(FC_Table_AB,aes(x=PosLog2FC, y=NegLog2FC))+ geom_point() + geom_vline(xintercept = 0) + 
  geom_hline(yintercept = 0) + theme_minimal() + ylim(-1, 1) + ggplot2::geom_smooth(method=lm, se=FALSE)  +
        stat_regline_equation() + stat_cor(method = "pearson", cor.coef.name = "R",label.x= -4, label.y=0.5) 

#ggsave("plot_AB_negpos.pdf", heigh=4,width=5)

```


### Gene Ontology
*Over Represented Analysis by GO*

```{r}
SigAnnot_AB_Pos <- SigAnnot_AB_Pos %>%
  add_column(Infection = "Pos")

SigAnnot_AB_Neg <- SigAnnot_AB_Neg %>%
  add_column(Infection = "Neg")

Compare_AB_NegPos = rbind (SigAnnot_AB_Pos,SigAnnot_AB_Neg)

Compare_AB_NegPos = Compare_AB_NegPos %>%
  mutate(Direction= case_when(log2FC < 0 ~ "Down",
                          log2FC > 0 ~ "UP",
                          )) 

Universe_AB_NegPos = c(annot_AB_Neg$ensembl_gene_id,annot_AB_Pos$ensembl_gene_id)
Universe_AB_NegPos = unique(Universe_AB_NegPos)

Compare_AB_NegPos_BP <- compareCluster(ensembl_gene_id~Direction+Infection, 
                                   data=Compare_AB_NegPos, 
                                   fun="enrichGO",
                                   OrgDb= org.Hs.eg.db,
                keyType       = 'ENSEMBL',
                universe = Universe_AB_NegPos,
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05)

dotplot(Compare_AB_NegPos_BP, x = 'Infection', showCategory=15,label_format=50, font=10) + facet_grid(~Direction)+ scale_colour_gradient(low = "#39c49b", high = "#41296b")
ggsave("ORA_AB_PosNeg.pdf", height=10)
Compare_AB_NegPos_BP_Annot <- setReadable(Compare_AB_NegPos_BP, 'org.Hs.eg.db', 'ENSEMBL')
write.csv(Compare_AB_NegPos_BP_Annot@compareClusterResult,"ORA_AB_PosNeg.csv")
```

GSEA
```{r}
AB_Pos_geneList=annot_AB_Pos$log2FC
names(AB_Pos_geneList) = as.character(annot_AB_Pos$ensembl_gene_id)
AB_Pos_geneList = na.omit(AB_Pos_geneList)
AB_Pos_geneList = sort(AB_Pos_geneList, decreasing = TRUE)

AB_Neg_geneList=annot_AB_Neg$log2FC
names(AB_Neg_geneList) = as.character(annot_AB_Neg$ensembl_gene_id)
AB_Neg_geneList = na.omit(AB_Neg_geneList)
AB_Neg_geneList = sort(AB_Neg_geneList, decreasing = TRUE)

inputList <- list(Pos = AB_Pos_geneList , Neg = AB_Neg_geneList)

Compare_AB_NegPos_BP_GSEA=compareCluster(geneClusters=inputList,  fun="gseGO",
                     OrgDb        = org.Hs.eg.db,
              ont          = "BP",
              eps = 1e-30,
              keyType       = 'ENSEMBL',
              minGSSize    = 10,
              maxGSSize    = 2000,
              pvalueCutoff = 0.05,
              pAdjustMethod="BH",
              verbose      = TRUE)

Compare_AB_NegPos_BP_GSEA_Annot <- setReadable(Compare_AB_NegPos_BP_GSEA, 'org.Hs.eg.db', 'ENSEMBL')
write.csv(Compare_AB_NegPos_BP_GSEA_Annot@compareClusterResult,"GSEA_AB_PosNeg.csv")

Compare_AB_NegPos_BP_GSEA <- pairwise_termsim(Compare_AB_NegPos_BP_GSEA)    

net <- emapplot(Compare_AB_NegPos_BP_GSEA, showCategory = 200,cex_line = 0.2, max.overlaps=10,group_category = T, group_legend = T,node_label = "group",nCluster=16,cex_category=4, nWords=5)
net

ggsave(plot=net,filename = "GSEA_AB_PosNeg_net_cluster.pdf", height=8, width=13)
```

