---
title: "CHAVA_Deconv"
author: "Carolina"
date: "2023-04-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ABsolute Immune Signal (ABIS) deconvolution
I ran ABsolute Immune Signal (ABIS) deconvolution on their shiny app https://giannimonaco.shinyapps.io/ABIS/

## B vs A
pos and Neg
```{r}
#Infection
ABIS_PosNeg_wide= subset(ABIS_absolute_wide, Category!="A2")
ABIS_PosNeg_long= subset(ABIS_absolute_long, Category!="A2")

#compare seropositive vs seroneg
ABIS_PosNeg_Ifxn_test = subset(ABIS_PosNeg_long, select=-c(Sex,Infected,Category))
ABIS_PosNeg_Ifxn_test =  ABIS_PosNeg_Ifxn_test%>%
  pivot_wider(
    names_from = c(CellType, InfectCat),
    values_from = c(Absolute))


ABIStest_PosNeg_Ifxn_abs=data.frame()

# stats
for (cell in colnames(ABIS_PosNeg_wide)[5:21]) {
  temp=wilcox.test(ABIS_PosNeg_Ifxn_test[[paste(cell,"Pos_A",sep="_")]], ABIS_PosNeg_Ifxn_test[[paste(cell,"Pos_B",sep="_")]],  alternative="two.sided", data=ABIS_PosNeg_Ifxn_test, exact=FALSE)
  ABIStest_PosNeg_Ifxn_abs=rbind(ABIStest_PosNeg_Ifxn_abs,c(paste(cell),c("AvB_pos"),temp$p.value))
  
  temp=wilcox.test(ABIS_PosNeg_Ifxn_test[[paste(cell,"Neg_A",sep="_")]], ABIS_PosNeg_Ifxn_test[[paste(cell,"Neg_B",sep="_")]],  alternative="two.sided", data=ABIS_PosNeg_Ifxn_test, exact=FALSE)
  ABIStest_PosNeg_Ifxn_abs=rbind(ABIStest_PosNeg_Ifxn_abs,c(paste(cell),c("AvB_neg"),temp$p.value))
  
  temp=wilcox.test(ABIS_PosNeg_Ifxn_test[[paste(cell,"Neg_A",sep="_")]], ABIS_PosNeg_Ifxn_test[[paste(cell,"Pos_A",sep="_")]],  alternative="two.sided", data=ABIS_PosNeg_Ifxn_test, exact=FALSE)
  ABIStest_PosNeg_Ifxn_abs=rbind(ABIStest_PosNeg_Ifxn_abs,c(paste(cell),c("-v+_A"),temp$p.value))
  
  temp=wilcox.test(ABIS_PosNeg_Ifxn_test[[paste(cell,"Neg_B",sep="_")]], ABIS_PosNeg_Ifxn_test[[paste(cell,"Pos_B",sep="_")]],  alternative="two.sided", data=ABIS_PosNeg_Ifxn_test, exact=FALSE)
  ABIStest_PosNeg_Ifxn_abs=rbind(ABIStest_PosNeg_Ifxn_abs,c(paste(cell),c("-v+_B"),temp$p.value))
}
  
colnames(ABIStest_PosNeg_Ifxn_abs)=c("CellType","Comparison", "Wilcox_P")
ABIStest_PosNeg_Ifxn_abs$FDR=p.adjust(ABIStest_PosNeg_Ifxn_abs$Wilcox_P, method = "BH", n = length(ABIStest_PosNeg_Ifxn_abs$Wilcox_P))
ABIStest_PosNeg_Ifxn_abs[order(ABIStest_PosNeg_Ifxn_abs$Wilcox_P),]
```

```{r}
ggplot(subset(ABIS_Pos_long,CellType =="T_CD8_n"|CellType =="T_CD4_n"|CellType =="T_CD4_Mem"|CellType =="T_CD8_mem"|CellType=="MAIT"|CellType =="B_n"|CellType =="B_mem"|CellType =="Monocytes_C"|CellType =="Monocytes_NC.I"|CellType =="mDCs"), aes(fill=Category,x=Category, y=Absolute)) +
  geom_boxplot(position="dodge", width=0.7, outlier.size = 1)+
 scale_fill_manual(breaks = c("A","B"),
                     values=c( "#0fbab7", "#6e016b")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

# CIBERSORT

I ran cybersortx on https://cibersortx.stanford.edu/. I used the LM22 microarray signature matrix. Since my data is from bulk RNAseq, I perform batch correction in B mode using LM22 source GEP file to account for the difference in microarray and RNAseqd data. I run 1000 permutations

```{r}
cyber_absolute_wide=read.csv("~/Library/CloudStorage/OneDrive-JohnsHopkins/Dr.GIlman/CHAVA_RNAseq/CIBERSORT/CHAVA_CIBERSORT_absolute.csv")
cyber_absolute_wide$Infected=as.factor(ifelse(cyber_absolute_wide$Infected==1,"Pos","Neg"))
cyber_absolute_wide$Sex=as.factor(ifelse(cyber_absolute_wide$Sex==1,"F","M"))
cyber_absolute_wide=na.omit(cyber_absolute_wide)
cyber_absolute_wide=cyber_absolute_wide[,1:26]

cyber_absolute_long = cyber_absolute_wide %>%
  pivot_longer(!c(Mixture,Category,Infected,Sex), names_to = "CellType", values_to = "Absolute")
```

## A v B

stats
```{r}
cyber_AB_wide= subset(cyber_absolute_wide, Category!="A2")
cyber_AB_long= subset(cyber_absolute_long, Category!="A2")

cyber_AB_long$InfectCat= paste(cyber_AB_long$Infected,cyber_AB_long$Category,sep = "_")


test_cyber_AB=compare_means(Absolute ~ InfectCat, data = cyber_AB_long, 
              group.by = "CellType", p.adjust.method="BH")

test_cyber_AB=test_cyber_AB[order(test_cyber_AB$p.format),]
test_cyber_AB
```


```{r}
ggplot(subset(cyber_AB_long,CellType =="B_n"|CellType =="B_mem"|CellType =="PC"|CellType =="Monocytes"|CellType =="DC_rest"|CellType =="DC_activ"|CellType =="T_CD8"|CellType =="T_CD4_n"|CellType =="T_CD4_mem_rest"|CellType =="T_CD4_mem_activ"|CellType =="T_CD4_n"|CellType =="NK_activ"|CellType =="NK_rest"), aes(fill=InfectCat,x=InfectCat, y=Absolute)) +
  geom_boxplot(position="dodge",width=0.7, outlier.size = 1) +
 scale_fill_manual(breaks = c("Pos_A","Pos_B","Neg_A","Neg_B"),
                   values=c( "#0fbab7", "#6e016b","#bbeefa","#edceec")) +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
  facet_wrap(~CellType, scales="free_y")

ggsave("AB_CIBERSORT.pdf", width=9.5, heigh=6.2)
```


# xCell


```{r}
library("immunedeconv")   
TPMcounts=read.delim("~/Library/CloudStorage/OneDrive-JohnsHopkins/Dr.GIlman/CHAVA_RNAseq/CIBERSORT/CHAVA_Cybersort_TPM.txt",row.names = 1)

#EPIC=immunedeconv::deconvolute(TPMcounts, "epic") #this one is missing a lot of immune cells
xcell=immunedeconv::deconvolute(TPMcounts, "xcell")

txcell<-as.data.frame(t(xcell[-1]))
colnames(txcell) <- xcell$cell_type

txcell=txcell[1:(length(txcell)-3)]

info <- CHAVAinfo %>%
   select(NovogeneCode, Sex, Infection, Cat)

txcell <- txcell %>% 
   rownames_to_column("NovogeneCode") %>% 
   left_join(info, by="NovogeneCode")

txcell=na.omit(txcell)
```

## A v B
```{r}
xcell_ab_wide= subset(txcell, Cat!="A2"&Cat!="NA")
xcell_ab_long= subset(txcell_long, Cat!="A2"&Cat!="NA")

xcell_ab_long$InfectCat=paste(xcell_ab_long$Infection,xcell_ab_long$Cat,sep = "_")


test_xcell_ab=compare_means(Absolute ~ InfectCat, data = xcell_ab_long, 
              group.by = "CellType", p.adjust.method="BH")

test_xcell_ab=test_xcell_ab[order(test_xcell_ab$p.format),]
test_xcell_ab
```

```{r}
ggplot(subset(xcell_ab_long,CellType =="Myeloid dendritic cell activated"|
CellType =="Monocyte"|CellType =="Myeloid dendritic cell"|CellType =="Macrophage"|CellType =="B cell memory"|CellType =="B cell"|CellType =="B cell plasma"|CellType =="B cell naive"|CellType =="T cell CD4+ naive"|CellType =="T cell CD8+ central memory"|CellType =="T cell CD8+ naive"|CellType =="T cell CD4+ memory"|CellType =="T cell CD8+"|CellType =="T cell gamma delta"|CellType =="NK cell"), aes(fill=InfectCat,x=InfectCat, y=Absolute)) +
  geom_boxplot(position="dodge",width=0.7, outlier.size = 1)+ 
 scale_fill_manual(breaks = c("Pos_A","Pos_B","Neg_A","Neg_B"),
                   values=c( "#0fbab7", "#6e016b","#bbeefa","#edceec")) +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
  facet_wrap(~CellType, scales="free_y")

ggsave("AB_xCell.pdf", width=9.5, heigh=7.5)

```
